---
title: "Matrices_analysis"
output: html_document
date: "2024-07-09"
editor_options: 
  chunk_output_type: inline
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Libraries

```{r}
library(ggplot2)
library(dplyr)
library(TADkitdev2)
library(BiocGenerics)
library(GenomicRanges)
library(devtools)
source("~/Documents/Stage projet/Stage_Gabriel/Fonctions/_R/Useful_functions.R")
```

# Parameters

```{r}
directory = "/home/gchartier/Documents/Stage/Orca_Thomas/newscreening/"
saveDir = "/home/gchartier/Documents/Stage/Results/"
resol = "32kb/"
output_type = "Obs"
chromsize.df = data.frame(chr = "1", size = 158534110)

mat_OE.path = (list.files(paste0(directory, resol), pattern = "orca_predictions_8Mb.txt", recursive = TRUE, full.names = TRUE)) %>% 
  .[order(nchar(.), .)]

mat_Obs.path = (list.files(paste0(directory, resol), pattern = "orca_normmats_8Mb.txt", recursive = TRUE, full.names = TRUE)) %>% 
  .[order(nchar(.), .)]

# Empty lists for individual scores
wtISlocal.df = data.frame()
mutPC1.lst = list()
compMUT.lst = list()
mutISglobal.lst = list()
mutISlocal.df = data.frame()
mutCompIS.df = data.frame()

# Empty lists for comparison scores
ratioIS.df = data.frame()
compRatio.df = data.frame()
SIC.df = data.frame()
localSIC.df = data.frame()
PC1corr.df = data.frame()
MATcorr.df = data.frame()

scores.df = data.frame()
```

# Load all matrices

## Load WT matrices

```{r}
# Obs / Exp
WT_OE.lst = readOrcaMat(
  path_predicted_mat = paste0(directory, "wildtype/orca_predictions_8Mb.txt")
)

# Obs
WT_Obs.lst = readOrcaMat(
  path_predicted_mat = paste0(directory, "wildtype/orca_predictions_8Mb.txt"),
  path_norm_mat = paste0(directory, "wildtype/orca_normmats_8Mb.txt"),
  output = "Obs"
)
```

## Load MUT matrices

```{r}
# Obs / Exp
MUT_OE.lst = lapply(mat_OE.path, readOrcaMat)
names(MUT_OE.lst) = paste0("MUT", 1:length(MUT_OE.lst), "_OE")

# Obs
MUT_Obs.lst = mapply(readOrcaMat, mat_OE.path, mat_Obs.path, output_type = output_type, SIMPLIFY = FALSE)
names(MUT_Obs.lst) = paste0("MUT", 1:length(MUT_Obs.lst), "_Obs")
```

# Calculate the different scores

## Calculate the WT individual scores

### WT PC1 and compartment calling

```{r}
##### PC1 #####

# Calculate the PC1 score for the WT
wtPC1.gr = matPCA(
  matrix = WT_OE.lst$Matrix,
  bin.width = WT_OE.lst$parameters$bin_width)

# Save the PC1 data in a .txt file as a data frame
# Convert the .gr tmp data into a .df
wtPC1.df = data.frame(wtPC1.gr) %>% cbind(., names = 1:nrow(.))

# Create the .txt file
write.table(wtPC1.df, paste0(saveDir, "wildtype/WT_PC1.txt"), row.names = FALSE, sep = "\t")

##### Compartments #####

# Call the compartments
compWT.gr = PC1calling(bedgraph = wtPC1.gr)

# Save the Comp data in a .txt file as a data frame
# Convert the .gr tmp data into a .df
compWT.df = data.frame(compWT.gr) %>% cbind(., names = 1:nrow(.))

# Create the .txt file
write.table(compWT.df, paste0(saveDir, "wildtype/WT_comp.txt"), row.names = FALSE, sep = "\t")
```

### WT global Insulation Score

```{r}
# Create a data frame with the values of the global IS
wtISglobal.df = globalInsScore(
  matrix = WT_Obs.lst$Matrix,
  start = WT_Obs.lst$parameters$start,
  stop = WT_Obs.lst$parameters$end,
  bin.width = WT_Obs.lst$parameters$bin_width,
  window = 128e3) %>% cbind(., names = 1:nrow(.))

# Global IS is converted to gr to plot a bedgraph later
wtISglobal.gr = dataframes2grange(wtISglobal.df, chromsize = chromsize.df, metadata.mcols = 4)

# Save the global IS data in a .txt file as a data frame
# Create the .txt file
write.table(wtISglobal.df, paste0(saveDir, "wildtype/WT_global_IS.txt"), row.names = FALSE, sep = "\t")

# Remove the data frame "wtISglobal.df" to save memory
rm(wtISglobal.df)
```

### WT local Insulation Score

```{r}
for (MAT in 1:length(MUT_OE.lst)) {
  IS = localInsScore(
    matrix = WT_Obs.lst$Matrix,
    view.point = (MUT_Obs.lst[[MAT]]$parameters$mutation_start + MUT_Obs.lst[[MAT]]$parameters$mutation_end) / 2,
    bin.width = MUT_Obs.lst[[MAT]]$parameters$bin_width,
    window = 128e3
  )
  tempIS.df = data.frame(chr = 1, mutation_start = MUT_Obs.lst[[MAT]]$parameters$mutation_start,
                             mutation_end = MUT_Obs.lst[[MAT]]$parameters$mutation_end, WT_local_IS = IS)
  wtISlocal.df = rbind(wtISlocal.df, tempIS.df)
  }
wtISlocal.df
```

## Calculate the MUT individual scores

```{r}
for (MAT in 1:length(MUT_OE.lst)) {
     
     
     ##### Calculate PC1 and call compartments #####
     
     
     # Calculate the PC1 and add it to a tmp
     tmp_PC1.gr = matPCA(
       matrix = MUT_OE.lst[[MAT]]$Matrix,
       bin.width = MUT_OE.lst[[MAT]]$parameters$bin_width)
     
     # Add the tmp to the list with all gr
     mutPC1.lst = append(mutPC1.lst, list(tmp_PC1.gr))
     
     # Save the PC1 data in a .txt file as a data frame
     # Convert the .gr tmp data into a .df
     tmp_PC1.df = data.frame(tmp_PC1.gr) %>% cbind(., names = 1:nrow(.))
     
     # Create the .txt file
     write.table(tmp_PC1.df, paste0(saveDir, resol, "PC1/", MAT, ".txt"), row.names = FALSE, sep = "\t")
     
     # Call the compartment and add it to a tmp
     tmp_comp.gr = PC1calling(bedgraph = tmp_PC1.gr)
     
     # Add the tmp to tha list with all gr
     compMUT.lst = append(compMUT.lst, list(tmp_comp.gr))
     
     # Save the Comp data in a .txt file as a data frame
     # Convert the .gr tmp data into a .df
     tmp_comp.df = data.frame(tmp_comp.gr) %>% cbind(., names = 1:nrow(.))
     
     # Create the .txt file
     write.table(tmp_comp.df, paste0(saveDir, resol, "Comp/", MAT, ".txt"), row.names = FALSE, sep = "\t")
     
     
     ##### Calculate the global Insulation Score #####
     
     
     # Calculate the global IS and add it to a tmp
     tmp_gIS.df = globalInsScore(
       matrix = MUT_Obs.lst[[MAT]]$Matrix,
       start = MUT_Obs.lst[[MAT]]$parameters$start,
       stop = MUT_Obs.lst[[MAT]]$parameters$end,
       bin.width = MUT_Obs.lst[[MAT]]$parameters$bin_width,
       window = 128e3) %>% cbind(., names = 1:nrow(.))
     
     # Global IS is converted to gr to plot a bedgraph later
     tmp_gIS.gr = dataframes2grange(tmp_gIS.df, chromsize = chromsize.df, metadata.mcols = 4)
     
     # Add the tmp to the list with all gr
     mutISglobal.lst = append(mutISglobal.lst, list(tmp_gIS.gr))
     
     # Save the global IS data in a .txt file as a data frame
     # Create the .txt file
     write.table(tmp_gIS.df, paste0(saveDir, resol, "global_IS/", MAT, ".txt"), row.names = FALSE, sep = "\t")
     
     # Remove the data frame "tmp_gIS.df" to save memory
     rm(tmp_gIS.df)
     
     
     ##### Calculate the local Insulation Score #####
     
     
     # Calculate the middle of the bin mutation
     bin.mid = (MUT_Obs.lst[[MAT]]$parameters$mutation_start + MUT_Obs.lst[[MAT]]$parameters$mutation_end) / 2
     
     # Calculate the local IS and add it to a tmp
     tmp_lIS.num = localInsScore(
       matrix = MUT_Obs.lst[[MAT]]$Matrix,
       view.point = bin.mid,
       bin.width = MUT_Obs.lst[[MAT]]$parameters$bin_width,
       window = 128e3)
     
     # Create a df tmp to store the main
     tmp_lIS.df = data.frame(chr = 1, mutation_start = MUT_Obs.lst[[MAT]]$parameters$mutation_start,
                             mutation_end = MUT_Obs.lst[[MAT]]$parameters$mutation_end, MUT_local_IS = tmp_lIS.num)
     
     # Add the tmp to the df with all num
     mutISlocal.df = rbind(mutISlocal.df, tmp_lIS.df)
     
     # Save the local IS data in a .txt file as a data frame
     # Create the .txt file
     write.table(mutISlocal.df, paste0(saveDir, resol, "local_IS.txt"), row.names = FALSE, sep = "\t")
     
     
     ##### Calculate the local Insulation Score with our boundary #####
     
     
     # Calculate the local IS and add it to a tmp
     tmp_compIS.num = localInsScore(
       matrix = MUT_Obs.lst[[MAT]]$Matrix,
       view.point = 3669000,
       bin.width = MUT_Obs.lst[[MAT]]$parameters$bin_width,
       window = 128e3)
     
     # Create a df tmp to store the main
     tmp_compIS.df = data.frame(chr = 1, mutation_start = MUT_Obs.lst[[MAT]]$parameters$mutation_start,
                             mutation_end = MUT_Obs.lst[[MAT]]$parameters$mutation_end, comp_local_IS = tmp_compIS.num)
     
     # Add the tmp to the df with all num
     mutCompIS.df = rbind(mutCompIS.df, tmp_compIS.df)
     
     # Save the local IS data in a .txt file as a data frame
     # Create the .txt file
     write.table(mutCompIS.df, paste0(saveDir, resol, "comp_IS.txt"), row.names = FALSE, sep = "\t")
     
     
     # Progress of the function
     print(paste0(MAT, " matrices done over ", length(MUT_OE.lst)))
}

# Give names for each elements in lists
names(mutPC1.lst) = paste0("MUT", 1:length(MUT_OE.lst), "_PC1")
names(compMUT.lst) = paste0("MUT", 1:length(MUT_OE.lst), "_compartments")
names(mutISglobal.lst) = paste0("MUT", 1:length(MUT_OE.lst), "_global_IS")

mutPC1.lst
compMUT.lst
mutISglobal.lst
mutISlocal.df
mutCompIS.df
```

## Calculate the comparison scores

```{r}
for (MAT in 1:length(MUT_OE.lst)) {
  
  #MAT = 2
  
  
  ##### Calculate the ratio between MUT local IS and WT local IS #####
  
  
  # Calculate the ratio
  ratio = (log2(mutISlocal.df$MUT_local_IS[MAT] / wtISlocal.df$WT_local_IS[MAT]))
  
  # Add the ratio to a tmp df
  tmp_ratioIS.df = data.frame(chr = 1, mutation_start = MUT_Obs.lst[[MAT]]$parameters$mutation_start,
                              mutation_end = MUT_Obs.lst[[MAT]]$parameters$mutation_end, ratio_IS = ratio)
  
  # Add the tmp to the df with all num
  ratioIS.df = rbind(ratioIS.df, tmp_ratioIS.df)
  
  # Save the ratio IS data in a .txt file as a data frame
  # Create the .txt file
  write.table(ratioIS.df, paste0(saveDir, resol, "ratio_IS.txt"), row.names = FALSE, sep = "\t")
  
  
  ##### Calculate the ratio between MUT local IS and WT local IS compared to our boundary #####
  
  
  # Calculate the ratio
  compRatio = (log2(mutCompIS.df$comp_local_IS[MAT] / wtISlocal.df$WT_local_IS[115]))
  
  # Add the ratio to a tmp df
  tmp_compRatio.df = data.frame(chr = 1, mutation_start = MUT_Obs.lst[[MAT]]$parameters$mutation_start,
                              mutation_end = MUT_Obs.lst[[MAT]]$parameters$mutation_end, comp_ratio = compRatio)
  
  # Add the tmp to the df with all num
  compRatio.df = rbind(compRatio.df, tmp_compRatio.df)
  
  # Save the ratio IS data in a .txt file as a data frame
  # Create the .txt file
  write.table(compRatio.df, paste0(saveDir, resol, "comp_ratio.txt"), row.names = FALSE, sep = "\t")
  
  
  ##### Calculate the Structural Impact Score of the mutation #####
  
  
  # Calculate the SIC
  sic.num = StructImpact(
    mutated.matrix = MUT_Obs.lst[[MAT]]$Matrix,
    wildtype.matrix = WT_Obs.lst$Matrix,
    mutation.start = MUT_Obs.lst[[MAT]]$parameters$mutation_start,
    bin.width = MUT_Obs.lst[[MAT]]$parameters$bin_width)
  
  # Add the SIC to a tmp df
  tmp_SIC.df = data.frame(chr = 1, mutation_start = MUT_Obs.lst[[MAT]]$parameters$mutation_start,
                              mutation_end = MUT_Obs.lst[[MAT]]$parameters$mutation_end, SIC = sic.num)
  
  # Add the tmp to the df with all num
  SIC.df = rbind(SIC.df, tmp_SIC.df)
  
  # Save the local SIC data in a .txt file as a data frame
  # Create the .txt file
  write.table(SIC.df, paste0(saveDir, resol, "SIC.txt"), row.names = FALSE, sep = "\t")
  
  
  ##### Calculate the local insulation Structural Impact Score of the mutation #####
  
  
  # Calculate the local SIC
  localSIC.num = localSIC(
    mutated.matrix = MUT_Obs.lst[[MAT]]$Matrix,
    wildtype.matrix = WT_Obs.lst$Matrix,
    mutation.start = MUT_Obs.lst[[MAT]]$parameters$mutation_start,
    bin.width = MUT_Obs.lst[[MAT]]$parameters$bin_width)
  
  # Add the SIC to a tmp df
  tmp_localSIC.df = data.frame(chr = 1, mutation_start = MUT_Obs.lst[[MAT]]$parameters$mutation_start,
                               mutation_end = MUT_Obs.lst[[MAT]]$parameters$mutation_end, local_SIC = localSIC.num)
  
  # Add the tmp to the df with all num
  localSIC.df = rbind(localSIC.df, tmp_localSIC.df)
  
  # Save the local SIC data in a .txt file as a data frame
  # Create the .txt file
  write.table(localSIC.df, paste0(saveDir, resol, "local_SIC.txt"), row.names = FALSE, sep = "\t")
  
  
  ##### Calculate the PC1 correlation #####
  
  
  # Calculate the PC1 correlation between MUT and WT
  PC1corr.bg = bgCorr(
    bedgraph.lst = list(WT = wtPC1.gr, MUT = mutPC1.lst[[MAT]]))
  
  # Add the PC1corr to a tmp df
  tmp_PC1corr.df = data.frame(chr = 1, mutation_start = MUT_Obs.lst[[MAT]]$parameters$mutation_start,
                              mutation_end = MUT_Obs.lst[[MAT]]$parameters$mutation_end, PC1corr = PC1corr.bg[2,1])
  
  # Add the tmp to the df with all num
  PC1corr.df = rbind(PC1corr.df, tmp_PC1corr.df)
  
  # Save the PC1corr data in a .txt file as a data frame
  # Create the .txt file
  write.table(PC1corr.df, paste0(saveDir, resol, "PC1corr.txt"), row.names = FALSE, sep = "\t")
  
  
  ##### Calculate the matrices correlation #####
  
  
  # Calculate the correlation between MUT and WT matrices
  MATcorr.bg = matCorr(
    matrix.lst = list(MUT_OE.lst[[MAT]]$Matrix, WT_OE.lst$Matrix))
  
  # Add the correlation to a tmp df
  tmp_MATcorr.df = data.frame(chr = 1, mutation_start = MUT_OE.lst[[MAT]]$parameters$mutation_start,
                              mutation_end = MUT_OE.lst[[MAT]]$parameters$mutation_end, MATcorr = MATcorr.bg[2,1])
  
  # Add the tmp to the df with all num
  MATcorr.df = rbind(MATcorr.df, tmp_MATcorr.df)
  
  # Save the MATcorr data in a .txt file as a data frame
  # Create the .txt file
  write.table(MATcorr.df, paste0(saveDir, resol, "MATcorr.txt"), row.names = FALSE, sep = "\t")
}

ratioIS.df
compRatio.df
SIC.df
localSIC.df
PC1corr.df
MATcorr.df
```

# Plot numeric scores

## Create a data frame merging all numeric scores above

```{r}
# Create the data frame
scores.df = merge(ratioIS.df, compRatio.df)
scores.df = merge(scores.df, SIC.df)
scores.df = merge(scores.df, localSIC.df)
scores.df = merge(scores.df, PC1corr.df)
scores.df = merge(scores.df, MATcorr.df)
arrange(scores.df, mutation_start)

# Create a .txt file as a save
write.table(scores.df, paste0(saveDir, resol, "scores.txt"), row.names = FALSE, sep = "\t")
```

## Plot numeric scores for all matrices

```{r}
plot.scores.df = data.frame(mutation_start = scores.df$mutation_start,
                            scores = c(scores.df$ratio_IS, scores.df$comp_ratio, scores.df$SIC, scores.df$local_SIC, scores.df$PC1corr, scores.df$MATcorr),
                            score_type = c(rep("ratioIS", nrow(scores.df)),
                                           rep("compRatio", nrow(scores.df)),
                                           rep("SIC", nrow(scores.df)),
                                           rep("localSIC", nrow(scores.df)),
                                           rep("MATcorr", nrow(scores.df)),
                                           rep("PC1corr", nrow(scores.df))))
scores.plot = ggplot(plot.scores.df, aes(mutation_start, scores, col = score_type)) +
  geom_step() +
  #facet_grid(score_type ~ ., scales = "free")
  facet_grid(~factor(score_type, levels = c("ratioIS", "compRatio", "SIC", "localSIC", "MATcorr", "PC1corr")) ~ ., scales = "free") +
  geom_vline(xintercept = 3669000, col = "red", linetype = "dotted") +
  geom_vline(xintercept = 2464000, col = "black", linetype = "dotted") +
  geom_vline(xintercept = 5888000, col = "black", linetype = "dotted")
scores.plot
```

# Plot other scores for some mutations

## PC1 and global IS

### Prepare the data

```{r}
# Load TAD location from a file for a specific BV
TAD.bed = read.table("/home/gchartier/Documents/Stage/Datas_tuto/tad3654_10kb.bed")

# Convert the TAD.bed data from df to gr
TAD_domains.gr = dataframes2grange(TAD.bed, chromsize.df)

# Put the gr into a list for later (TAD plot)
TAD.lst = list(BV_3654 = TAD_domains.gr)

# Create a list with all the global IS we want to plot
analysed_IS.lst = list(WT = wtISglobal.gr, MUT78 = mutISglobal.lst$MUT78_global_IS, MUT115 = mutISglobal.lst$MUT115_global_IS, MUT185 = mutISglobal.lst$MUT185_global_IS)

# Create a list with all the PC1 we want to plot
analysed_PC1.lst = list(WT = wtPC1.gr, MUT78 = mutPC1.lst$MUT78_PC1, MUT115 = mutPC1.lst$MUT115_PC1, MUT185 = mutPC1.lst$MUT185_PC1)

# Create a list with all the comp we want to plot
analysed_comp.lst = list(MUT78 = compMUT.lst$MUT78_compartments, MUT115 = compMUT.lst$MUT115_compartments, MUT185 = compMUT.lst$MUT185_compartments)
```

### Plot the data

```{r}
# Plot for all the region
TAD.plot = mTADplot(
  tad.lst = TAD.lst,
  start = WT_OE.lst$parameters$start,
  stop = WT_OE.lst$parameters$end,
  chr = 1,
  bedgraph.lst = list(IS = analysed_IS.lst, PC1 = analysed_PC1.lst)
)

# Plot around the location we study
zoom1TAD.plot = mTADplot(
  tad.lst = TAD.lst,
  start = 1e6,
  stop = 7e6,
  chr = 1,
  bedgraph.lst = list(IS = analysed_IS.lst, PC1 = analysed_PC1.lst)
)

TAD.plot
zoom1TAD.plot
```

## PC1 correlation

### Calculate the Pearson correlation and the r^2

```{r}
# Calculate the Pearson correlation coefficient
PC1.pearson.lst = lapply(1:length(analysed_PC1.lst), function(i) {
  PC1.pearson.corr = cor.test(wtPC1.gr$PC1, analysed_PC1.lst[[i]]$PC1, method = "pearson", conf.level = 0.95, parse = TRUE)
  print(PC1.pearson.corr$estimate)
})

# Calculate the r^2
PC1.r2.lst = lapply(1:length(analysed_PC1.lst), function(i) {
  PC1.r2.coef = summary(lm(analysed_PC1.lst[[i]]$PC1 ~ wtPC1.gr$PC1))$r.squared
  print(PC1.r2.coef)
})
```

### Plot the PC1 correlation

```{r}
# Scatterplot with MUTi VS WT, Pearson correlation coefficient and r^2
PC1corr.plot = lapply(1:length(analysed_PC1.lst), function(i) {
  PC1corr.plot = ggplot(data.frame(analysed_PC1.lst[[i]]), aes(x = wtPC1.gr$PC1, y = analysed_PC1.lst[[i]]$PC1)) +
  geom_point(na.rm = TRUE) +
  labs(
    x = "WT",
    y = names(analysed_PC1.lst[i]),
    title = paste0(names(analysed_PC1.lst[i]), " and WT PC1 correlation")) +
    geom_smooth(method = "lm", na.rm = TRUE) +  # Trend line following a linear model method (lm)
    annotate("text", x = -0.04, y = 0.07, label = paste0("Pearson correlation = ", PC1.pearson.lst[[i]], "\n r² = ", PC1.r2.lst[[i]]), size = 3.2) +
  theme_bw()
})
PC1corr.plot
```

# Plot matrices for some mutations

## Create a list with analysed matrices

```{r}
# List with Obs matrices
analysed_Obs_MAT.lst = list(MUT78 = MUT_Obs.lst$MUT78_Obs, MUT115 = MUT_Obs.lst$MUT115_Obs, MUT185 = MUT_Obs.lst$MUT185_Obs)

# List with Obs / Exp matrices
analysed_OE_MAT.lst = list(MUT78 = MUT_OE.lst$MUT78_OE, MUT115 = MUT_OE.lst$MUT115_OE, MUT185 = MUT_OE.lst$MUT185_OE)
```

### Plot the correlation between ratio IS and local SIC

```{r}
# Scatterplot with MUTi VS WT, Pearson correlation coefficient and r^2
scoresCorr.plot = ggplot(scores.df, aes(x = scores.df$ratio_IS, y = scores.df$local_SIC)) +
  geom_point(na.rm = TRUE) +
  labs(
    x = "ratio IS",
    y = "local SIC",
    title = "Correlation between ratio IS and local SIC") +
    geom_smooth(method = "lm", na.rm = TRUE) +  # Trend line following a linear model method (lm)
    #annotate("text", x = -0.05, y = 0.05, label = paste0("Pearson correlation = ", PC1.pearson.lst[[i]], "\n r² = ", PC1.r2.lst[[i]]), size = 3.2) +
  theme_bw()
scoresCorr.plot
```

## MUTi & WT Obs matrices

### From start to stop

```{r}
MAT_Obs.lst = lapply(1:length(analysed_Obs_MAT.lst), function(MAT) {
  matObs = mMATplot(
    matrix.upper = analysed_Obs_MAT.lst[[MAT]]$Matrix,
    matrix.lower = WT_Obs.lst$Matrix,
    start = (analysed_Obs_MAT.lst[[MAT]]$parameters$start),
    stop = (analysed_Obs_MAT.lst[[MAT]]$parameters$end),
    bin.width = analysed_Obs_MAT.lst[[MAT]]$parameters$bin_width,
    matrix.upper.txt = "MUT",
    matrix.lower.txt = "WT",
    scale.colors = "H",
    tad.upper.line = analysed_comp.lst[[MAT]],
    tad.lower.line = compWT.gr) +
    ggtitle(paste0(names(analysed_Obs_MAT.lst[MAT]), " & WT Obs matrix"),
            subtitle = paste0("Mutation from ", analysed_Obs_MAT.lst[[MAT]]$parameters$mutation_start, " Mb to ",
                              analysed_Obs_MAT.lst[[MAT]]$parameters$mutation_end, " Mb")) +
    geom_hline(yintercept = -analysed_Obs_MAT.lst[[MAT]]$parameters$mutation_start, col = "black", linetype = "dotted") +  # Location of the mutation
    geom_hline(yintercept = -analysed_Obs_MAT.lst[[MAT]]$parameters$mutation_end, col = "black", linetype = "dotted") +
    geom_vline(xintercept = analysed_Obs_MAT.lst[[MAT]]$parameters$mutation_start, col = "black", linetype = "dotted") +
    geom_vline(xintercept = analysed_Obs_MAT.lst[[MAT]]$parameters$mutation_end, col = "black", linetype = "dotted") +
    geom_hline(yintercept = -3669000, col = "darkgoldenrod4", linetype = "dotted") +  # Location of the compartment edge
    geom_vline(xintercept = 3669000, col = "darkgoldenrod4", linetype = "dotted")
})
MAT_Obs.lst
```

### Zoom: from -3Mb to +3Mb from mutation

```{r}
zoomMAT_Obs.lst = lapply(1:(length(analysed_Obs_MAT.lst)), function(MAT) {
  matObs = mMATplot(
    matrix.upper = analysed_Obs_MAT.lst[[MAT]]$Matrix,
    matrix.lower = WT_Obs.lst$Matrix,
    start = (analysed_Obs_MAT.lst[[MAT]]$parameters$mutation_start) - 2e6,
    stop = (analysed_Obs_MAT.lst[[MAT]]$parameters$mutation_end) + 2e6,
    bin.width = analysed_Obs_MAT.lst[[MAT]]$parameters$bin_width,
    matrix.upper.txt = "MUT",
    matrix.lower.txt = "WT",
    scale.colors = "H",
    tad.upper.line = analysed_comp.lst[[MAT]],
    tad.lower.line = compWT.gr) +
    ggtitle(paste0("Zoom ", names(analysed_Obs_MAT.lst[MAT]), " & WT Obs matrix"),
            subtitle = paste0("Mutation from ", analysed_Obs_MAT.lst[[MAT]]$parameters$mutation_start, " Mb to ",
                              analysed_Obs_MAT.lst[[MAT]]$parameters$mutation_end, " Mb")) +
    geom_hline(yintercept = -analysed_Obs_MAT.lst[[MAT]]$parameters$mutation_start, col = "black", linetype = "dotted") +  # Location of the mutation
    geom_hline(yintercept = -analysed_Obs_MAT.lst[[MAT]]$parameters$mutation_end, col = "black", linetype = "dotted") +
    geom_vline(xintercept = analysed_Obs_MAT.lst[[MAT]]$parameters$mutation_start, col = "black", linetype = "dotted") +
    geom_vline(xintercept = analysed_Obs_MAT.lst[[MAT]]$parameters$mutation_end, col = "black", linetype = "dotted") +
    geom_hline(yintercept = -3669000, col = "darkgoldenrod4", linetype = "dotted") +  # Location of the compartment edge
    geom_vline(xintercept = 3669000, col = "darkgoldenrod4", linetype = "dotted")
})
zoomMAT_Obs.lst
```

## MUTi & WT Obs / Exp matrices

### From start to stop

```{r}
MAT_OE.lst = lapply(1:length(analysed_OE_MAT.lst), function(MAT) {
  matOE = mMATplot(
    matrix.upper = analysed_OE_MAT.lst[[MAT]]$Matrix,
    matrix.lower = WT_OE.lst$Matrix,
    start = (analysed_OE_MAT.lst[[MAT]]$parameters$start),
    stop = (analysed_OE_MAT.lst[[MAT]]$parameters$end),
    bin.width = analysed_OE_MAT.lst[[MAT]]$parameters$bin_width,
    matrix.upper.txt = "MUT",
    matrix.lower.txt = "WT",
    scale.colors = "OE",
    tad.upper.line = analysed_comp.lst[[MAT]],
    tad.lower.line = compWT.gr) +
    ggtitle(paste0(names(analysed_OE_MAT.lst[MAT]), " & WT Obs / Exp matrix"),
            subtitle = paste0("Mutation from ", analysed_OE_MAT.lst[[MAT]]$parameters$mutation_start, " Mb to ",
                              analysed_OE_MAT.lst[[MAT]]$parameters$mutation_end, " Mb")) +
    geom_hline(yintercept = -analysed_OE_MAT.lst[[MAT]]$parameters$mutation_start, col = "black", linetype = "dotted") +  # Location of the mutation
    geom_hline(yintercept = -analysed_OE_MAT.lst[[MAT]]$parameters$mutation_end, col = "black", linetype = "dotted") +
    geom_vline(xintercept = analysed_OE_MAT.lst[[MAT]]$parameters$mutation_start, col = "black", linetype = "dotted") +
    geom_vline(xintercept = analysed_OE_MAT.lst[[MAT]]$parameters$mutation_end, col = "black", linetype = "dotted") +
    geom_hline(yintercept = -3669000, col = "darkgoldenrod4", linetype = "dotted") +  # Location of the compartment edge
    geom_vline(xintercept = 3669000, col = "darkgoldenrod4", linetype = "dotted")
})
MAT_OE.lst
```

### Zoom: from -3Mb to +3Mb from mutation

```{r}
zoomMAT_OE.lst = lapply(1:(length(analysed_OE_MAT.lst)), function(MAT) {
  matOE = mMATplot(
    matrix.upper = analysed_OE_MAT.lst[[MAT]]$Matrix,
    matrix.lower = WT_OE.lst$Matrix,
    start = (analysed_OE_MAT.lst[[MAT]]$parameters$mutation_start - analysed_OE_MAT.lst[[MAT]]$parameters$start) - 2e6,
    stop = (analysed_OE_MAT.lst[[MAT]]$parameters$mutation_end - analysed_OE_MAT.lst[[MAT]]$parameters$start) + 2e6,
    bin.width = analysed_OE_MAT.lst[[MAT]]$parameters$bin_width,
    matrix.upper.txt = "MUT",
    matrix.lower.txt = "WT",
    scale.colors = "OE",
    tad.upper.line = analysed_comp.lst[[MAT]],
    tad.lower.line = compWT.gr) +
    ggtitle(paste0("Zoom ", names(analysed_OE_MAT.lst[MAT]), " & WT Obs / Exp matrix"),
            subtitle = paste0("Mutation from ", analysed_OE_MAT.lst[[MAT]]$parameters$mutation_start, " Mb to ",
                              analysed_OE_MAT.lst[[MAT]]$parameters$mutation_end, " Mb")) +
    geom_hline(yintercept = -analysed_OE_MAT.lst[[MAT]]$parameters$mutation_start, col = "black", linetype = "dotted") +  # Location of the mutation
    geom_hline(yintercept = -analysed_OE_MAT.lst[[MAT]]$parameters$mutation_end, col = "black", linetype = "dotted") +
    geom_vline(xintercept = analysed_OE_MAT.lst[[MAT]]$parameters$mutation_start, col = "black", linetype = "dotted") +
    geom_vline(xintercept = analysed_OE_MAT.lst[[MAT]]$parameters$mutation_end, col = "black", linetype = "dotted") +
    geom_hline(yintercept = -3669000, col = "darkgoldenrod4", linetype = "dotted") +  # Location of the compartment edge
    geom_vline(xintercept = 3669000, col = "darkgoldenrod4", linetype = "dotted")
})
zoomMAT_OE.lst
```

## Muti & WT Fold Change

### From start to stop

```{r}
MAT_FC.lst = lapply(1:length(analysed_Obs_MAT.lst), function(MAT) {
  matFC = MATplot(
    matrix = analysed_Obs_MAT.lst[[MAT]]$Matrix / WT_Obs.lst$Matrix,
    start = analysed_Obs_MAT.lst[[MAT]]$parameters$start,
    stop = analysed_Obs_MAT.lst[[MAT]]$parameters$end,
    bin.width = analysed_Obs_MAT.lst[[MAT]]$parameters$bin_width,
    scale.colors = "OE2",
    tad.upper.line = analysed_comp.lst[[MAT]],
    tad.lower.line = analysed_comp.lst[[MAT]]) +
    ggtitle(paste0(names(analysed_Obs_MAT.lst[MAT]), " & WT Fold change matrix"),
            subtitle = paste0("Mutation from ", analysed_Obs_MAT.lst[[MAT]]$parameters$mutation_start, " Mb to ",
                              analysed_Obs_MAT.lst[[MAT]]$parameters$mutation_end, " Mb")) +
    geom_hline(yintercept = -analysed_Obs_MAT.lst[[MAT]]$parameters$mutation_start, col = "black", linetype = "dotted") +  # Location of the mutation
    geom_hline(yintercept = -analysed_Obs_MAT.lst[[MAT]]$parameters$mutation_end, col = "black", linetype = "dotted") +
    geom_vline(xintercept = analysed_Obs_MAT.lst[[MAT]]$parameters$mutation_start, col = "black", linetype = "dotted") +
    geom_vline(xintercept = analysed_Obs_MAT.lst[[MAT]]$parameters$mutation_end, col = "black", linetype = "dotted") +
    geom_hline(yintercept = -3669000, col = "darkgoldenrod4", linetype = "dotted") +  # Location of the compartment edge
    geom_vline(xintercept = 3669000, col = "darkgoldenrod4", linetype = "dotted")
})
MAT_FC.lst
```

### Zoom: from -3Mb to +3Mb from mutation

```{r}
zoomMAT_FC.lst = lapply(1:(length(analysed_Obs_MAT.lst)), function(MAT) {
  matFC = MATplot(
    matrix = analysed_Obs_MAT.lst[[MAT]]$Matrix / WT_Obs.lst$Matrix,
    start = analysed_Obs_MAT.lst[[MAT]]$parameters$mutation_start - 2e6,
    stop = analysed_Obs_MAT.lst[[MAT]]$parameters$mutation_end + 2e6,
    bin.width = analysed_Obs_MAT.lst[[MAT]]$parameters$bin_width,
    scale.colors = "OE2",
    tad.upper.line = analysed_comp.lst[[MAT]],
    tad.lower.line = analysed_comp.lst[[MAT]]) +
    ggtitle(paste0("Zoom ", names(analysed_Obs_MAT.lst[MAT]), " & WT Fold change matrix"),
            subtitle = paste0("Mutation from ", analysed_Obs_MAT.lst[[MAT]]$parameters$mutation_start, " Mb to ",
                              analysed_Obs_MAT.lst[[MAT]]$parameters$mutation_end, " Mb")) +
    geom_hline(yintercept = -analysed_Obs_MAT.lst[[MAT]]$parameters$mutation_start, col = "black", linetype = "dotted") +  # Location of the mutation
    geom_hline(yintercept = -analysed_Obs_MAT.lst[[MAT]]$parameters$mutation_end, col = "black", linetype = "dotted") +
    geom_vline(xintercept = analysed_Obs_MAT.lst[[MAT]]$parameters$mutation_start, col = "black", linetype = "dotted") +
    geom_vline(xintercept = analysed_Obs_MAT.lst[[MAT]]$parameters$mutation_end, col = "black", linetype = "dotted") +
    geom_hline(yintercept = -3669000, col = "darkgoldenrod4", linetype = "dotted") +  # Location of the compartment edge
    geom_vline(xintercept = 3669000, col = "darkgoldenrod4", linetype = "dotted")
})
zoomMAT_FC.lst
```

